---
title: "RLadies August Meeting: Functions"
author: "Ellen Fitzsimmons & Hope Snyder"    
date: "8/25/20"
header-includes:
  - \usepackage{amsmath, amssymb,amsthm,amsfonts}
  - \usepackage{multicol}
  - \usepackage[usenames,dvipsnames]{pstricks}
  - \usepackage{graphicx}
  - \usetheme{Warsaw}
  - \usecolortheme{dolphin}
  - \AtBeginSection{}
  - \AtBeginSubsection{}
output: 
  beamer_presentation:
    toc: false
    slide_level: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Basics

## What is a Function?

### What is a function?

- A function is a set of statements organized together to perform a certain task

\[y= mx+b\]

## Basic parts

### Basic Parts
- Name
- Input
- Computation
- Output

## Example
### Computing the Slope

Name:

```{r, eval = FALSE, echo = TRUE}
slope <- function(){}
```

### Computing the Slope

Input:

```{r, eval = FALSE, echo = TRUE}
slope <- function(X, Y, B){}
```

```{r, eval = FALSE, echo = TRUE}
slope <- function(coords, constant){}
```

### Computing the Slope

Computations:

```{r, eval = FALSE, echo = TRUE}
slope <- function(X, Y, B){
  (Y-B)/X
}
```

### Computing the Slope

Output:

```{r, eval = FALSE, echo = TRUE}
slope <- function(X, Y, B){
  M = (Y-B)/X
  
  print("The slope is equal to ", M)
}
```

### Computing the Slope

Use:

```{r, echo = TRUE}
slope <- function(X, Y, B){
  M = (Y-B)/X
  
  print(paste("The slope is equal to ", M))
}

slope(2,5,7)
```

# Advanced

## When to Use

### When to Use

- Almost whenever you want!
- When your code repeats a lot
- Something you will want to use in the future
- When you want to apply the code to multiple things at once
- Writing a package

## Nesting functions

### Nesting functions

Functions can be called from within other functions!

```{r, echo = TRUE}
dat <- mtcars

pick_a_color <- function(x){
  I=length(unique(x))
  myColor=hsv((1:I)/I,1,.7,1)
  return(myColor)
}

My_plot=function(x,y,...){
  color = pick_a_color(x)
  
  boxplot(y~x,col=color,... )
}
```

### Nesting functions
```{r, echo = TRUE, fig.height = 7, fig.width = 12}
My_plot(dat$gear,dat$mpg,xlab="gear", ylab="MPG")
```


## Using Functions on Multiple Values

### `purrr` 

You can apply functions to a table of values!

```{r, echo = TRUE}
lm(dat$mpg ~ dat$gear)
```

### `purrr` 

```{r, echo = TRUE}
slope <- function(Y,X){
  (Y-5.623)/X
}

library(purrr)
mat <- purrr::map2_dbl(dat$mpg, dat$gear, slope)
```

### `purrr` 

```{r, echo = TRUE}
mat <- matrix(mat, ncol = 4, byrow = TRUE)
mat
```






